% ------------------------------------------------
\StartAcknowledgments
% ------------------------------------------------

I would like to express my sincere gratitude to all those who have supported and guided me throughout this research journey.

First and foremost, I extend my deepest appreciation to my advisor, Professor Cheng-Te Li, for his invaluable guidance, patience, and continuous support throughout my master's program. His expertise in graph neural networks and fake news detection provided the foundation for this research, and his constructive feedback helped shape this work into its current form. Professor Li's dedication to research excellence and his commitment to fostering innovative thinking have been instrumental in my academic development.

I am grateful to the members of my thesis committee for their time, expertise, and thoughtful feedback during the thesis defense. Their insightful questions and suggestions have significantly improved the quality and rigor of this work.

I would like to acknowledge the contributions of the research community that developed the foundational technologies used in this thesis. The FakeNewsNet dataset creators provided high-quality benchmarks essential for evaluating fake news detection methods. The developers of DeBERTa, PyTorch Geometric, and other open-source tools enabled the implementation of our complex heterogeneous graph neural network framework.

Special thanks to my fellow graduate students and research group members for their collaboration, discussions, and moral support. The weekly lab meetings and informal research discussions provided valuable opportunities to refine ideas and overcome technical challenges. Their diverse perspectives and constructive criticism helped improve both the technical content and presentation of this work.

I am deeply grateful to my family for their unwavering support and encouragement throughout my academic journey. Their understanding during long hours of research and writing, and their belief in my abilities, provided the motivation necessary to complete this challenging project.

I also acknowledge the financial support provided by the National Science and Technology Council and National Cheng Kung University, which made this research possible and allowed me to focus on my studies without financial burden.

Finally, I want to thank the broader fake news detection research community for their commitment to combating misinformation. The importance of this research area has become increasingly evident as digital misinformation poses growing threats to democratic societies and public health. I hope this work contributes meaningfully to the ongoing efforts to develop more effective and practical solutions for detecting and countering false information.

The completion of this thesis represents not only a personal academic milestone but also a small contribution to the critical challenge of maintaining information integrity in our digital age. I am honored to have had the opportunity to work on such an important problem and hope that future researchers will build upon these foundations to create even more effective solutions.

% ------------------------------------------------
\EndAcknowledgments
% ------------------------------------------------
